Impact of engines scheduler on the performance of Big Data pipelines in HPC

### 2021-01-21 latex/paper.tex
Finished first writeo of the engine section

\subsection{Dask}
Dask is a Python-based Big Data engine with growing popularity in the scientific Python ecosystem.
To mitigate the data transfer bottleneck in Big Data workflows, Dask was designed with data locality and in-memory in mind.
Data locality, popularized by Map-Reduce \cite{dean2008mapreduce}, schedules tasks where the data reside.
In-memory computing minimize the overhead of transfering data to disk by keeping data in memory when possible.
Dask uses lazy evalutation to reduce unnecessary communication and computation.
The engine builds a dynamic graph prior to execution allowing to determine which task to compute.
Dask workflows can further reduce data transfer by using leveraging mulithreading whenever the application is not bounded by Python's GIL.
Fault-tolerance is achieved by recording of data lineage, the sequence of operations used on to modify the initial data.

Dask offers five data structures:
\href{https://docs.dask.org/en/latest/array.html}{Array},
\href{https://docs.dask.org/en/latest/bag.html}{Bag},
\href{https://docs.dask.org/en/latest/dataframe.html}{DataFrame},
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed},
and \href{https://docs.dask.org/en/latest/futures.html}{Futures}.
The Dask Array offers a clone of NumPy API for distributed processing of large arrays.
A Dask Bag is a distributed collection of Python object that offers a programming abstraction similar to \href{https://toolz.readthedocs.io/en/latest/}{PyToolz}.
Dask Dataframe is a parallel composition of Pandas Dataframes used to process large amount of tabular data.
Dask Delayed offers an API for distributing abrirtary function that do not fit in the above frameworks.
Lastly, Dask Futures can also execute abrirtary functions, however, it launch computation immediately rather than lazily.
Dask modularity allow users to install only required components making it lightweigth.

In Dask a scheduler decides where and when to execute tasks using the Dask graph.
API operations generate multiple fine-coarsed tasks in the computation graph allwing an easier representatin of complex algorithms.

The Dask engine is compatible to numerous distributed schedulers, including YARN and Mesos.
Dask also provide its own \textit{Dask Distributed scheduler}.
Althought Dask is compatible with schedulers available in HPC clusters,
We chose to use Dask Distributed scheduler, to keep the environment balanced between the engines.

In Dask Distributed scheduler, a \textit{dask-scheduler} process admistrates the resource provided by \textit{dask-worker} in the cluster.
The scheduler receives jobs from clients and assigns tasks to available workers.
Dask scheduler use a FIFO (First-In-First-Out) job scheduling policy.
That is, completely process a branch of the Dask graph before proceding with the next one.

Dask offers multiple ways to deploy a cluster including: with SSH configs, Kubernetes, SLURM, PBS, and others.
For our experiments, we used the \href{https://jobqueue.dask.org/en/latest/generated/dask_jobqueue.SLURMCluster.html}{Dask SLURM cluster} API.

@@

\subsection{Apache Spark}
Apache Spark is a widely-used general pupose Big Data engine.
Like Dask, it aims at reducing data transfer cost by incorporating data locality and in-memory computing, and lazy evaluation.

Spark primary data structure is Resilient Distributed Dataset (RDD)\cite{RDD}, a fault-tolerant, parallel collection of data elements.
RDDs is the basis of the other Spark data structure: Datasets and DataFrames.
Datasets are similar to RDD but benefits from additional performance by leveraging the Spark SQL's optimized execution engine.
The DataFrame API are Datasets organized into named-columns and used to process tabular data.
While the DataFrame API is available in all supported language, DAtasets are limited to Sacla and Java.

Python is a common programming language in the scientific community.
This is particulary true in the neuroimaging field with numerous specialized Python libraries to operate on data.
While serialization from Python to Java create overhead, we found it to be minimal \cite{8943502}.
We chose to focus on PySpark API to have a more balance environment between the different engines and due to its suitability to neuroimaging.

>>>>

\subsection{Dask}
Dask is a Python-based Big Data engine with growing popularity in the scientific Python ecosystem.
Dask was designed with data locality and in-memory in mind to mitigate the data transfer bottleneck in Big Data workflows.
Data locality, popularized by Map-Reduce \cite{dean2008mapreduce}, schedules tasks where the data reside.
In-memory computing minimizes the overhead of transferring data to disk by keeping data in memory when possible.
Dask uses lazy evaluation to reduce unnecessary communication and computation.
The engine builds a dynamic graph before execution, allowing it to determine which task to compute.
Dask workflows can further reduce data transfer by leveraging multithreading whenever Python's GIL does not restrict it.
Fault-tolerance is achieved by recording data lineage: the sequence of operations used to modify the initial data.

Dask offers five data structures:
\href{https://docs.dask.org/en/latest/array.html}{Array},
\href{https://docs.dask.org/en/latest/bag.html}{Bag},
\href{https://docs.dask.org/en/latest/dataframe.html}{DataFrame},
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed},
and \href{https://docs.dask.org/en/latest/futures.html}{Futures}.
Arrays offer a clone of NumPy API for distributed processing of large arrays.
Bags are a distributed collection of Python object that offers a programming abstraction similar to \href{https://toolz.readthedocs.io/en/latest/}{PyToolz}.
Dataframes are a parallel composition of \href{https://pandas.pydata.org/}{Pandas} Dataframes used to process a large amount of tabular data.
Dask Delayed offers an API for distributing arbitrary functions that do not fit in the above frameworks.
Lastly, Dask Futures can also execute arbitrary functions; however, it launches computation immediately rather than lazily.
Dask modularity allows users to install only required components making it lightweight.

In Dask, a scheduler decides where and when to execute tasks using the Dask graph.
API operations generate multiple fine-coarse tasks in the computation graph, allowing a more straightforward representation of complex algorithms.

The Dask engine is compatible with multiple distributed schedulers, including YARN and Mesos.
Dask also provides its own \textit{Dask Distributed scheduler}.
We chose to use Dask Distributed scheduler to keep the environment balanced between the engines.

In the Dask Distributed scheduler, a \textit{dask-scheduler} process administrates the resource provided by  \textit{dask-worker}s in the cluster.
The scheduler receives jobs from clients and assigns tasks to available workers.
Task scheduler uses a FIFO (First-In-First-Out) job scheduling policy.
That is an utter process branch of the Dask graph before proceeding with the next one.

Dask offers multiple ways to deploy a cluster, including, but not limited to, SSH configs, Kubernetes, SLURM, PBS.
For our experiments, we used the \href{https://jobqueue.dask.org/en/latest/generated/dask_jobqueue.SLURMCluster.html}{Dask SLURM cluster} API.

@@

\subsection{Apache Spark}
Apache Spark is a widely-used general-purpose Big Data engine.
Like Dask, it aims at reducing data transfer costs by incorporating data locality, in-memory computing, and lazy evaluation.

Spark offers three options to schedule jobs: Spark Standalone, Mesos, and YARN.
The Spark standalone is a simple built-in scheduler.
YARN is mainly used to schedule Hadoop-based workflows, while Mesos can be used for various workflows.
We limit our focus to Spark Standalone scheduler, as researchers are likely to execute their workflows in an HPC environment where, usually, neither YARN nor Mesos is available.

In the Spark Standalone scheduler, a \textit{leader} coordinates the resource provisioned by \textit{workers} in the cluster.
A \textit{driver}  process receives jobs from clients and requests workers from the leader.
Jobs are divided into stages to be executed onto workers.
Each operation in a stage is represented by a high-level task in the computation graph.
Like Dask, Spark Standalone scheduler uses a FIFO policy to schedule tasks.
Spark Standalone has two execution modes: (1) the client mode, where the driver process runs in a dedicated process,
and (2) the cluster mode, where the driver runs within a worker process.
Our experiments use the client mode since cluster mode is not available in PySpark.

Spark's primary data structure is Resilient Distributed Dataset (RDD)\cite{RDD}, a fault-tolerant, parallel collection of data elements.
RDDs are the basis of the other Spark data structure: Datasets and DataFrames.
Datasets are similar to RDD but benefit additional performance by leveraging the Spark SQL's optimized execution engine. 
The DataFrames are Datasets organized into named-columns and are used to process tabular data. 
While the DataFrame API is available in all supported languages, Datasets are limited to Scala and Java.

Python is a standard programming language in the scientific community.
This is particularly true in the neuroimaging field, with numerous specialized Python libraries to operate on data.
While serialization from Python to Java creates overhead, we found it to be minimal \cite{8943502}.
We focus on PySpark API to have a more balanced environment between the different engines and for its suitability to neuroimaging.
